---
title: "Practice with t-tests"
output: html_document
date: "2022-10-27"
---

## Reviewing the t-test

Let's conduct a 2-sample t-test to compare body fat percentage between men and women.

The data set presents the body fat percentage for a sample of men and women who all work out at the same gym. In general, men are leaner than women, with a body fat percentage centering around 15% whereas for women it tends to center around 20%. Let’s see if the same is true of these data.

## Instructions

Use the information you learned from reading chapter 5 and doing homework 21 to conduct a t-test comparing mean body fat percentage between the male and female groups. But don’t just rush to the test – follow the recommended analysis workflow:

1.  Plot the data and look at expected relationships

Load data
```{r}
rm(list = ls())
library(tidyverse)
library(here)
bodyfat <- read.csv(here("Data/body_fat_percentage.csv"))

bodyfat$gender <- 
  as.factor(bodyfat$gender)
levels(bodyfat$gender)
```

Plot
```{r}
ggplot(bodyfat, aes(y = body_fat_percent, x = gender)) +
  geom_boxplot()
```

2.  Calculate the mean and standard deviation for the data and examine the resulting output.
```{r}
bodyfat |> 
  group_by(gender) |>
  summarise(mean = mean(body_fat_percent, na.rm = T),
            sd = sd(body_fat_percent, na.rm = T))
```

3.  Discuss with your partner:
    -   Review in the book: What are the assumptions about the data appropriate for a t-test? When you examine the plots and summary data, do they appear to meet those assumptions?
equal variance, normality, residuals 
    -   Based on examination of the plots and summary data, do you expect that there is a statistically significant difference in mean body fat percentage between males and females in this sample?
4.  Conduct a t-test and then interpret the results. What do they tell you about the null vs. alternative hypotheses?
```{r}
t.test(body_fat_percent ~ gender, data = bodyfat)
```

